---
title: "Eurobarometer analysis"
output:
  md_document:
    variant: gfm
---

Data source: [Eurobarometer 98.2 (2023)](https://search.gesis.org/research_data/ZA7953)

```{r setup, include=FALSE}
library(tidyverse)
```

## Read data

From the data source above, download the file `ZA7953_v1-0-0_eurobarometer_98.2.dta` and extract the columns for country names and answers to the question `sd18a - DEMOCRACY SATISFACTION - COUNTRY` ("On the whole, are you very satisfied, fairly satisfied, not very satisfied or not at all satisfied with the way democracy works in (OUR COUNTRY)?")


```{r cars}
d = haven::read_dta("../../data/raw/ZA7953_v1-0-0_eurobarometer_98.2.dta") %>% 
  select("isocntry", "sd18a") %>%
  mutate(isocntry = gsub("DE-.","DE", isocntry)) #merge east and west germany

#extract metadata
meta = data.frame(values = attributes(d$sd18a)$labels %>% unname,
             labels = attributes(d$sd18a)$labels %>% names)

```

## Summarize and clean data

The raw data contains answers from every single participant. For analysis, summarize by country and answer. Append metadata and country names.

```{r}
#calculate numbers and shares of answers
d = d %>%
  group_by(isocntry, sd18a) %>%
  summarise(n = n(), .groups = "drop") %>% 
  mutate(share = n/sum(n)) %>% 
  left_join(meta, by = c("sd18a" = "values")) %>% #merge descriptions
  left_join(read.csv("eurobarometer_countrycodes.csv"), by = join_by(isocntry)) #merge country names
```



## Analyze democracy satisfaction

Make dataset of democracy satisfaction shares by country and save. Print overall satisfaction levels.

```{r}
#make dataset by country
x = d %>% 
  select(-n, -sd18a) %>% 
  pivot_wider(names_from = labels, values_from = share) %>%
  mutate("Satisfied TOTAL" = `Very satisfied`+`Fairly satisfied`,
         "Not satisfied TOTAL" = `Not very satisfied` + `Not at all satisfied`)

#save data
write.csv(x, file = "../../data/processed/eurobarometer_country_democracy_satisfaction.csv", row.names = F)

#overall satisfaction
d %>% group_by(sd18a, labels) %>% 
  summarise(n = sum(n), .groups = "drop") %>% ungroup %>% mutate(share = n / sum(n)) %>% 
  knitr::kable()
```
